---
title: "621_Homework 4"
output:
  pdf_document:
    toc: true
    toc_depth: 3
author: "Liting Hu"
---
```{r global_options, include=FALSE}
knitr::opts_chunk$set(warning=FALSE, message=FALSE)
```

## Question 1
### 1
```{r}
require(Sim.DiffProc)
setwd("/Users/apple/Desktop/621/HW4")
sample_data <- read.csv("sample_data.csv") # Read data
sample_data <- ts(sample_data, deltat=1/365) # dt = 1/365
head(sample_data)

```

```{r}
model.match <- as.data.frame(matrix(nrow = 5, ncol = 2))
model.match[, 1] <- colnames(sample_data)
best.n <- c()
for (i in 1:5) {
    # model 1
    fx1 <- expression( theta[1]*x )
    gx1 <- expression( theta[2]*x^theta[3] )
    mod1 <- fitsde(data=sample_data[, i], drift=fx1, diffusion=gx1,
                   start = list(theta1=1, theta2=1, theta3=1), pmle="euler")
    
    # model 2
    fx2 <- expression( theta[1]+theta[2]*x )
    gx2 <- expression( theta[3]*x^theta[4] )
    mod2 <- fitsde(data=sample_data[, i], drift=fx2, diffusion=gx2,
                   start = list(theta1=1,theta2=1,theta3=1,theta4=1), pmle="euler")
    
    # model 3
    fx3 <- expression( theta[1]+theta[2]*x )
    gx3 <- expression( theta[3]*sqrt(x) )
    mod3 <- fitsde(data=sample_data[, i], drift=fx3, diffusion=gx3,
                   start = list(theta1=1,theta2=1,theta3=1), pmle="euler")
    
    # model 4
    fx4 <- expression( theta[1] )
    gx4 <- expression( theta[2]*x^theta[3] )
    mod4 <- fitsde(data = sample_data[, i], drift=fx4, diffusion=gx4,
                   start = list(theta1=1,theta2=1,theta3=1), pmle="euler")
    
    # model 5
    fx5 <- expression( theta[1]*x )
    gx5 <- expression( theta[2]+theta[3]*x^theta[4] )
    mod5 <- fitsde(data=sample_data[, i], drift=fx5, diffusion=gx5,
                   start = list(theta1=1,theta2=1,theta3=1,theta4=1), pmle="euler")
    
    # Computes AIC
    AIC <- c(AIC(mod1),AIC(mod2),AIC(mod3),AIC(mod4),AIC(mod5))
    k <- which.min(AIC) # Choose the minimum
    best.n[i] <- k
    best <- paste0("model", k)
    model.match[i, 2] <- best
}
```

```{r results = 'asis'}
knitr::kable(model.match, caption = "Model match")
```
The best models are chosen as Model 1,1,1,5,1 respectively based on AIC.

### 2
```{r}
fx <- list(fx1, fx2, fx3, fx4, fx5)
gx <- list(gx1, gx2, gx3, gx4, gx5)
pmle <- eval(formals(fitsde.default)$pmle)

for (j in 1:5) {
    fitres <- lapply(1:4, function(i) fitsde(data = sample_data[, j], drift=fx[[best.n[j]]], diffusion=gx[[best.n[j]]],
                                             pmle=pmle[i], start=list(theta1=1,theta2=1,theta3=1,theta4=1)))
    Coef <- data.frame(do.call("cbind", lapply(1:4, function(i) coef(fitres[[i]]))))
    Info <- data.frame(do.call("rbind", lapply(1:4, function(i) logLik(fitres[[i]]))),
                       do.call("rbind", lapply(1:4, function(i) AIC(fitres[[i]]))),
                       do.call("rbind", lapply(1:4, function(i) BIC(fitres[[i]]))))
    Info <- t(Info)
    colnames(Coef) <- pmle
    colnames(Info) <- pmle
    rownames(Info) <- c("logLic", "AIC", "BIC")
    result <- rbind(Coef, Info)
    print(paste0("stock", j))
    print(result)
}
```

For stock 1, the minimum AIC and BIC appear in euler method. Those calculated by kessler method are much larger (8 >> -2.563912e+05, 23.0258709 >> -2.563762e+05).

For stock 2, the minimum AIC and BIC appear in kessler method. And euler's AIC and BIC are also less than ozaki's and shoji's.

For stock 3, the minimum AICs and BICs appear in euler and kessler methods which are much smaller than ozaki's and shoji's.

For stock 4, the minimum AIC and BIC appear in euler method. And kessler's AIC and BIC are also less than ozaki's and shoji's.

For stock 5, the minimum AIC and BIC appear in euler method. And kessler's AIC and BIC are also less than ozaki's and shoji's. (similar situation as stock 4)

### 3
For my opinion, the Euler method and Kessler method provide best estimates.

## Question 2
```{r}
library(xlsx)
swaptions <- read.xlsx("2017_2_15_mid.xlsx", sheetName = "Quotes") # Read data
head(swaptions)

column <- 5 # choose 5th maturity which is 5 years
dim(swaptions)
Vol <- swaptions[seq(1, 38, by=2), 7]/100 # At-the-money volatility
Strike <- swaptions[seq(2, 38, by=2), 7]/100 # Strike price K
expiry <- swaptions[seq(1, 38, by=2), 1] # expiration date

temp <- gregexpr("[0-9]+", expiry)
expiry <- as.numeric(unlist(regmatches(expiry, temp)))
expiry[1:4] <- expiry[1:4]/12 # Expiry date for swaptions
```


### 1
```{r}
SABRVol<-function(alpha,beta,rho,nu,Tm,f){ 
    # All the contracts are at the money, so f=K
        Term1 <- alpha/f^(1-beta)
        Term2 <- (1-beta)^2/24*alpha^2/f^(2-2*beta)+rho*beta*nu*alpha/4/f^(1-beta)+(2-3*rho^2)*nu^2/24
        ans <- Term1*(1 + Term2*Tm)
        return(ans)
}

optimization <- function(beta) {
    init.values = c(0.1, 0, 0.1)
    lower.bound = c(0.0001, -1, 0.0001) # -1 < rou < 1
    upper.bound = c(Inf, 1, Inf)
    
    objective.f <- function(parm,beta) {
        # Objective function
        # to search the minimum sum of (sigma_mkt - sigma_B)^2 
        alpha <- parm[1]
        rho <- parm[2]
        nu <- parm[3]
        Sigma_B <- SABRVol(alpha,beta,rho,nu,expiry,Strike)
        return(sum((Vol-Sigma_B)^2))
    }
    
    # PORT optimization routine
    opt<-nlminb(start=init.values,
                objective = objective.f,
                lower = lower.bound,
                upper = upper.bound,
                beta=beta)
    
    parms <-opt$par
    obj <- opt$objective
    ans <- t(as.data.frame(c(beta, parms, obj)))
    colnames(ans) <- c("beta", "alpha", "rho", "nu", "objective")
    rownames(ans) <- NULL
    return(ans)
    
}
re1 <- optimization(0.5)
```

```{r results = 'asis'}
knitr::kable(re1, caption = "Beta=0.5")
```

### 2

```{r}
re2 <- optimization(0.7)
re3 <- optimization(0.4)
```

```{r results = 'asis'}
knitr::kable(re2, caption = "Beta=0.7")
```

```{r results = 'asis'}
knitr::kable(re3, caption = "Beta=0.4")
```

### 3
```{r}
library(ggplot2)
library(reshape2)
res <- as.data.frame(rbind(re1, re2, re3))

res.long <- melt(res, id="beta")
ggplot(data=res.long, aes(x=beta, y=value, color=variable)) + geom_line()
```
From this plot, we can see that as beta (3 values) grows larger, the objective function decreases. While in the mean time, alpha and rho is getting larger. Nu is decreasing.

### 4
```{r}
betas <- seq(0, 1, by=0.1)
opts <- optimization(0)
for (i in betas) {
    a <- optimization(i)
    opts <- rbind(opts, a)
}

opts <- as.data.frame(opts[2:12, ])
```

```{r results = 'asis'}
knitr::kable(opts)
```

```{r}
opts.long <- melt(opts, id="beta")
ggplot(data=opts.long, aes(x=beta, y=value, color=variable)) + geom_line()
```

To be more precisely, in this section, let beta equal to 0, 0.1, 0.2, ... 1. Still, objective value is decreasing as beta growing. We can basically say that beta=1 gives me the best estimation. 

### 5

```{r}
beta <- 1
re5 <- optimization(beta)
alpha <- re3[2]
rho <- re3[3]
nu <- re3[4]

column <- 7 # choose 7th maturity which is 10 years
Vol2 <- swaptions[seq(1, 38, by=2), 9]/100 # At-the-money volatility
Strike2 <- swaptions[seq(2, 38, by=2), 9]/100 # Strike price K

VolB <- SABRVol(alpha,beta,rho,nu,expiry,Strike2)
plot(expiry, Vol2, type = "l", col = "red", ylab = "volitility")
lines(expiry, VolB, col = "blue")
legend("topright", legend=c("Benchmark", "Calculated vol"),
       col=c("red", "blue"), lty=c(1,1))
obj <- sum((VolB-Vol2)^2)
obj
```

Apply beta = 1 to swaps whose maturity is 10 years. The volitility calculated by SABR model is nearly a straight line. Given volitilities fluctuate around it. The SSE is 0.05352116 which is fairly small.

## 3

```{r}

S0=1
Tm=1
sigma=0.2
r=0.0075

alpha <- 1
N <- 10000
eta <- 50
lambda <- 2*pi/N/eta
a <- N*eta # the effective upper limit for the integration
b <- N*lambda/2 # log of strike prices range from -b to b

# for phi_T
f_phi<- function(u, s, mu, sigma) {
    # integrand
    ans <- 1/sqrt(2*pi*sigma^2)*exp(-(s-mu)^2/2/sigma^2)*exp(1i*u*s)
    return(ans)
}

phi_u <- function(u, mu, sigma, a, N) {
    # Integral
    # By the Simpson’s rule
    s <- seq(-a, a, length.out = N)
    h <- s[2] - s[1]
    wn <- rep(c(2*h/3, 4*h/3), floor(N/2))
    wn[1] <- h/3
    wn[N] <- h/3
    fn <- sapply(s, f_phi, u = u, mu = mu, sigma = sigma)
    ans <- sum(fn*wn)
    return(ans)
}

# Out the money
zeta_v <- function(v, r, Tm, mu, sigma, a, N) {
    ans <- exp(-r*Tm)*(1/(1 + 1i*v) - exp(r*Tm)/(1i*v) -
                           phi_u(v - 1i, mu, sigma, a, N)/(v^2 - 1i*v))
    return(ans)
}

gamma_v <- function(v, r, Tm, mu, sigma, alpha, a, N) {
    ans <- (zeta_v(v - 1i*alpha, r, Tm, mu, sigma, a, N)-
                zeta_v(v + 1i*alpha, r, Tm, mu, sigma, a, N))/2
    return(ans)
}

# In the money
psi_v <- function(v, r, Tm, mu, sigma, alpha, a, N) {
    ans <- (exp(-r*Tm)*phi_u(v - (alpha + 1)*1i, mu, sigma, a, N))/
        (alpha^2 + alpha - v^2 + 1i*(2*alpha + 1)*v)
    return(ans)
}


Call_FFT <- function(u, r, Tm, mu, sigma, alpha, a, N) {
    ku <- -b + lambda*(u-1)
    s <- seq(0.0001, a, length.out = N)
    h <- s[2] - s[1]
    wn <- rep(c(2*h/3, 4*h/3), floor(N/2))
    wn[1] <- h/3
    wn[N] <- h/3
    j <- seq(1, N)
    vj <- (j-1)*eta
    if (u < ((N+1)/2)) {
        f <- exp(-1i*2*pi/N*(j-1)*(u-1))*exp(1i*b*vj)*psi_v(vj, r, Tm, r, sigma, alpha, a, N)
        C <- exp(-alpha*k)/pi*sum(f*wn)
    } else {
        f <- exp(-1i*2*pi/N*(j-1)*(u-1))*exp(1i*b*vj)*gamma_v(vj, r, Tm, r, sigma, alpha, a, N)
        C <- sinh(alpha*k)/pi*sum(f*wn)
    }
    return(C)
}
u=60
Call_FFT(u, r, Tm, r, sigma, alpha, a, N)

K <- (-b + (u-1)*lambda)^2
Option_BSM <- function(isCall = T, S0=1, K=1, Tm=1, sigma=0.2, r=0.0075, div=0) {
    d1 <- (log(S0/K) + (r - div + sigma^2/2)*Tm)/sigma/sqrt(Tm)
    d2 <- d1 - sigma*sqrt(Tm)
    if (isCall) {p <- S0*exp(-div*Tm)*pnorm(d1) - K*exp(-r*Tm)*pnorm(d2)}
    else {p <- K*exp(-r*Tm)*pnorm(-d2) - S0*exp(-div*Tm)*pnorm(-d1)}
    return(p)
}
Option_BSM(K=K)

```

