---
title: "621_Homework 1"
output:
  pdf_document:
    toc: true
    toc_depth: 3
author: "Liting Hu"
---
```{r message = FALSE, echo = FALSE}
library(quantmod)
library(ggplot2)
library(rgl)
```

## Question 1
### a)

To calculate the European Call and Put Option, define functions "Callprice" and "Putprice" based on the solution of Black-Scholes formula.
```{r}
Callprice <- function(S0, tau, K, r, sigma) {
    d1 <- (log(S0/K)+(r-sigma^2/2)*tau)/tau/sqrt(tau)
    d2 <- d1 - sigma*sqrt(tau)
    c <- S0*pnorm(d1) - K*exp(-r*tau)*pnorm(d2)
    return(c)
}

Putprice <- function(S0, tau, K, r, sigma) {
    d1 <- (log(S0/K)+(r-sigma^2/2)*tau)/tau/sqrt(tau)
    d2 <- d1 - sigma*sqrt(tau)
    p <- K*exp(-r*tau)*pnorm(-d2) - S0*pnorm(-d1)
    return(p)
}

S0 <- 100       # Stock price
tau <- 30/252   # Time to maturity
K <- 100        # Strike price
r <- 0.05       # Interest rate
sigma <- 0.2    # Volatility
Callprice(S0, tau, K, r, sigma)
Putprice(S0, tau, K, r, sigma)
```
Under the given initial conditions, call option price is 3.0496 and the put option price is 2.4561.

### b)

```{r}
Callprice(S0, tau, K, r, sigma) - Putprice(S0, tau, K, r, sigma) - S0 + K*exp(-r*tau)
```
The Put-Call parity relation holds.

### c)
Get option chains from Yahoo finance by "quantmod":
```{r}
AAPL.OPTS <- getOptionChain("AAPL", NULL)
C1 <- AAPL.OPTS$Mar.17.2017$calls
C1$Ave.Price <- (C1$Bid + C1$Ask)/2
C1 <- C1[, c(1, 8)]
C2 <- AAPL.OPTS$Apr.21.2017$calls
C2$Ave.Price <- (C2$Bid + C2$Ask)/2
C2 <- C2[, c(1, 8)]
C3 <- AAPL.OPTS$Jul.21.2017$calls
C3$Ave.Price <- (C3$Bid + C3$Ask)/2
C3 <- C3[, c(1, 8)]

temp <- merge(C1, C2, by = "Strike")
calls <- merge(temp, C3, by = "Strike")
colnames(calls) <- c("Strike", "Mar.17.2017", "Apr.21.2017", "Oct.20.2017")
calls <- calls[1:20, ]
```

All call option prices are showed below:

```{r results = 'asis'}
knitr::kable(calls, caption = "Call option prices")
```


Get the actual stock price:
```{r}
todaystock <- getQuote("AAPL") 
S_0 <- todaystock[, 2]
S_0
```
The actual stock price is 135.72.

Treasury bill rate is 0.005. (https://www.treasury.gov/resource-center/data-chart-center/interest-rates/Pages/TextView.aspx?data=billrates)

\(f(\sigma)=C_{BSM}-C_M\) is defined as "fsigma":
```{r}
r <- 0.005
fsigma <- function(sigma, K_i, maturity_i) {
    cc <- calls[K_i, maturity_i + 1]
    K <- calls[K_i, 1]
    if (maturity_i == 1) tau = 23/252 # time to maturity
    else if(maturity_i == 2) tau = 48/252
    else tau = 111/252
    ans <- Callprice(S_0, tau, K, r, sigma) - cc
    return(ans)
}
```

We want to find a interval \([a, a + d]\)(\(d\) should be small enough) that makes the secant method converge
```{r}
interval <- calls
interval[, 2:4] <- NaN
delta <- 0.1
for(i in 1:20) {
    for(j in 1:3) {
        a <- seq(1, 5, by = delta)
        for(k in a) {
            if (fsigma(k,i,j)*fsigma(k + delta, i, j) < 0) {interval[i,j + 1] <- k}
        }
    }
}
```

```{r results = 'asis'}
knitr::kable(interval, caption = "Left side of intervals")
```
Here come NaN values which means there is no root for \(\sigma\) in such a conditon.

```{r}
ImpliedVolatility <- calls
ptm <- proc.time()
count <- 0
for(i in 1:20) {
    for(j in 1:3) {
        a <- interval[i, j + 1]
        if (is.na(a) == T) {
            ImpliedVolatility[i, j + 1] <- NaN
            next
            }
        b <- a + delta
        epsilon <- abs(a - b)
        while(epsilon > 1e-4) {
            count <- count +1
            mid <- (a + b)/2
            if(fsigma(a, i, j)*fsigma(mid, i, j) < 0 ) b <- mid
            else a <- mid
            epsilon <- abs(a - b)
        }
        ImpliedVolatility[i, j+1] <- a
    }
}
count
ImpliedVolatility
proc.time() - ptm
```

### d)
```{r}
# Secant
ImpliedVolatility2 <- calls
ptm <- proc.time()
count2 <- 0
for(i in 1:20) {
    for(j in 1:3) {
        x1 <- interval[i, j + 1]
        if (is.na(x1) == T) {
            ImpliedVolatility2[i, j + 1] <- NaN
            next
        }
        x2 <- x1 + delta
        dfxn <- (fsigma(x1, i, j) - fsigma(x2, i, j))/(x1 - x2)
        tang <- fsigma(x2, i, j)/dfxn
        epsilon <- abs(tang)
        while(epsilon > 1e-4) {
            count2 <- count2 + 1
            x1 <- x2
            x2 <- x2 - tang
            dfxn <- (fsigma(x1, i, j) - fsigma(x2, i, j))/(x1 - x2)
            tang <- fsigma(x2, i, j)/dfxn
            epsilon <- abs(tang)
        }
        ImpliedVolatility2[i, j+1] <- x2
    }
}
count2
ImpliedVolatility2
proc.time() - ptm
```
Both the steps and time spent on secant methods (130, 0.041) is less than those spent on bisection method (540, 0.078). I suppose the reason may be the interval is small enough and the secant method's order of convergence is 1.618 which is superlinear.

Also I found that if we set the initial intervals too large, the secant method seldom converge, although the bisection method always converge if there exist roots. So it is hard for me to compare these two methods when the initial intervals is large. (like [1, 10])

### e)
```{r warning = F}
temp1 <- as.matrix(ImpliedVolatility)
temp2 <- rbind(temp1[, 1:2], temp1[, c(1, 3)], temp1[, c(1, 4)])
maturity <- c(rep("Mar.17.2017", 20), rep("Apr.21.2017", 20), rep("Oct.20.2017", 20))
newdf <- data.frame(temp2, maturity)
colnames(newdf)[2] <- "Implied_Volatility"

ggplot(data = newdf, aes(x = Strike, y = Implied_Volatility, colour = maturity)) + geom_point()

```

All call option prices are getting smaller as the Strikes increasing. Longer time to maturity means lower decreasing speed of implied volatility.

Around the actual stock price the implied volatility tend to be equal under different maturities. 

This 3d plot cannot be presented in Rmarkdown pdf, but it works in R file.
```{r}
plot3d(ImpliedVolatility$Strike, 1, ImpliedVolatility$Mar.17.2017, col = "blue", ylim = c(.5, 3.5))
points3d(ImpliedVolatility$Strike, 2, ImpliedVolatility$Apr.21.2017, col = "green")
points3d(ImpliedVolatility$Strike, 3, ImpliedVolatility$Oct.20.2017, col = "red")
```

### f)
To use 3 methods (forward, backword and central) to evaluate delta, vega and gamma, define functions:
```{r}
Delta1 <- function(d, S0, tau, K, r, sigma, ty) {
    if (ty == "b") D <- (Callprice(S0, tau, K, r, sigma) - Callprice(S0 - d, tau, K, r, sigma))/d
    else if (ty == "f") D <- (Callprice(S0 + d, tau, K, r, sigma) - Callprice(S0, tau, K, r, sigma))/d
    else if (ty == "c") D <- (Callprice(S0 + d, tau, K, r, sigma) - Callprice(S0 - d, tau, K, r, sigma))/d/2
    else D <- NaN
    return(D)
}

Vega1 <- function(d, S0, tau, K, r, sigma, ty) {
    if (ty == "b") V <- (Callprice(S0, tau, K, r, sigma) - Callprice(S0, tau, K, r, sigma - d))/d
    else if (ty == "f") V <- (Callprice(S0, tau, K, r, sigma + d) - Callprice(S0, tau, K, r, sigma))/d
    else if (ty == "c") V <- (Callprice(S0, tau, K, r, sigma + d) - Callprice(S0, tau, K, r, sigma - d))/d/2
    else V <- NaN
    return(V)
}

Gamma1 <- function(d, S0, tau, K, r, sigma, ty) {
    if (ty == "b") {G <- (Callprice(S0, tau, K, r, sigma) - 2*Callprice(S0 - d, tau, K, r, sigma) + 
                              Callprice(S0 - 2*d, tau, K, r, sigma))/d^2}
    else if (ty == "f") {G <- (Callprice(S0 + 2*d, tau, K, r, sigma) - 2*Callprice(S0 + d, tau, K, r, sigma) + 
                                   Callprice(S0, tau, K, r, sigma))/d^2}
    else if (ty == "c") {G <- (Callprice(S0 + d, tau, K, r, sigma) - 2*Callprice(S0, tau, K, r, sigma) + 
                                       Callprice(S0 - d, tau, K, r, sigma))/d^2}
    else G <- NaN
    return(G)
}
```

For delta
```{r}
d1 <- seq(1, .01,by=-0.01)
fd_f <- sapply(d1, Delta1, S0=S0, tau=tau, K=K, r=r, sigma=sigma, "f")
fd_b <- sapply(d1, Delta1, S0=S0, tau=tau, K=K, r=r, sigma=sigma, "b")
fd_c <- sapply(d1, Delta1, S0=S0, tau=tau, K=K, r=r, sigma=sigma, "c")

type <- c(rep("forward", length(d1)), rep("backword", length(d1)), rep("central", length(d1)))

delta_dd <- data.frame("d" = rep(d1, 3), "delta" = c(fd_f, fd_b, fd_c), "type" = type)
ggplot(data = delta_dd, aes(x = d, y = delta, colour = type)) + geom_line()
```

For vega
```{r}
fv_f <- sapply(d1, Vega1, S0=S0, tau=tau, K=K, r=r, sigma=sigma, "f")
fv_b <- sapply(d1, Vega1, S0=S0, tau=tau, K=K, r=r, sigma=sigma, "b")
fv_c <- sapply(d1, Vega1, S0=S0, tau=tau, K=K, r=r, sigma=sigma, "c")

vega_dd <- data.frame("d" = rep(d1, 3), "vega" = c(fv_f, fv_b, fv_c), "type" = type)
ggplot(data = vega_dd, aes(x = d, y = vega, colour = type)) + geom_line()
```

For gamma
```{r}
fg_f <- sapply(d1, Gamma1, S0=S0, tau=tau, K=K, r=r, sigma=sigma, "f")
fg_b <- sapply(d1, Gamma1, S0=S0, tau=tau, K=K, r=r, sigma=sigma, "b")
fg_c <- sapply(d1, Gamma1, S0=S0, tau=tau, K=K, r=r, sigma=sigma, "c")

gamma_dd <- data.frame("d" = rep(d1, 3), "gamma" = c(fg_f, fg_b, fg_c), "type" = type)
ggplot(data = gamma_dd, aes(x = d, y = gamma, colour = type)) + geom_line()
```
These plots shows that all three parameters converge as \(d\) is approaching to \(0\).

```{r}
d <- 1e-5 # let d be small enough
Delta1(d, S0, tau, K, r, sigma, "f") #forward
Delta1(d, S0, tau, K, r, sigma, "b") #backward
Delta1(d, S0, tau, K, r, sigma, "c") #central

Vega1(d, S0, tau, K, r, sigma, "f") #forward
Vega1(d, S0, tau, K, r, sigma, "b") #backword
Vega1(d, S0, tau, K, r, sigma, "c") #central

Gamma1(d, S0, tau, K, r, sigma, "f") #forward
Gamma1(d, S0, tau, K, r, sigma, "b") #backword
Gamma1(d, S0, tau, K, r, sigma, "c") #central
```

So in initial conditon, \( delta = 0.5405\), \( vega = 13.5322\), \( gamma = 0.0320\).

### g)
If we apply the implied volatilities, access to the same process as f):

```{r}
delta_df <- ImpliedVolatility
vega_df <- ImpliedVolatility
gamma_df <- ImpliedVolatility

for (i in 1:20) {
    for (j in 1:3) {
        sigma <- ImpliedVolatility[i, j+1]
        if (is.na(sigma) == T) {
            delta_df[i, j+1] <- NaN
            vega_df[i, j+1] <- NaN
            gamma_df[i, j+1] <- NaN
            break}
        delta_df[i, j+1] <- Delta1(d, S_0, tau, K, r, sigma, "c")
        vega_df[i, j+1] <- Vega1(d, S_0, tau, K, r, sigma, "c")
        gamma_df[i, j+1] <- Gamma1(d, S_0, tau, K, r, sigma, "c")
    }
}
```

The results are showed below:

```{r results = 'asis'}
knitr::kable(delta_df, caption = "Delta")
```
```{r results = 'asis'}
knitr::kable(vega_df, caption = "Vega")
```
```{r results = 'asis'}
knitr::kable(gamma_df, caption = "Gamma")
```


## Question 2
### a)

Define the the real–valued function:
```{r}
fx <- function(x) {
    if (x == 0) fx <- 1
    else fx <- sin(x)/x
    return(fx)
}
```

For the trapezoidal rule:
```{r}
trapezoidal <- function(n, a) {
    h <- 2*a/(n-1)
    x <- seq(-a, a, by = h)
    wn <- rep(h, n)
    wn[1] <- h/2
    wn[n] <- h/2
    fn <- sapply(x, fx)
    ans <- sum(fn*wn)
    return(ans)
}
trapezoidal(1e6, 1e6)
```

For the Simpson’s quadrature rule

```{r}
simpson <- function(n, a) {
    h <- 2*a/(n-1)
    x <- seq(-a, a, by = h)
    wn <- rep(c(0, 4*h/3), floor(n/2)) + rep(c(2*h/3, 0), floor(n/2))
    wn[1] <- h/3
    wn[n] <- h/3
    fn <- sapply(x, fx)
    ans <- sum(fn*wn)
    return(ans)
}
simpson(1e6, 1e6)
```
Both the results are close to \(\pi\).

### b)
To compute the truncation errors, define functions "TE_trape" and "TE_simp" respectively:
```{r}
TE_trape <- function(n, a) {
    temp <- trapezoidal(n, a)
    te <- temp - pi
    return(te)
}

TE_simp <- function(n, a) {
    temp <- simpson(n, a)
    te <- temp - pi
    return(te)
}
```

(1) Fix \(a = 10^4\)
```{r}
# Truncation error of trapezoidal rule
a <- 1e4
N <- seq(100, 1e4, by = 45)
Truncation_error <- sapply(N, TE_trape, a=a) 
plot(N, Truncation_error, type = "l", col = "blue", main = "Truncation_error by trapezoidal rule")
```

```{r}
# Truncation error of Simpson's rule
a <- 1e4
N <- seq(100, 1e4, by = 45)
Truncation_error <- sapply(N, TE_simp, a=a) 
plot(N, Truncation_error, type = "l", col = "red", main = "Truncation_error by simpson's rule")
```

For both rules, the truncation error decrease progressively as n gets larger and larger. Besides, the simpson's rule converge later and fluctuate more.

(2) fix \(N = 10^6\):
```{r}
# Truncation error of trapezoidal rule
N <- 1e4
a <- seq(1e2, 1e4, by = 45)
Truncation_error <- sapply(a, TE_trape, n = N) 
plot(a, Truncation_error, type = "l", col = "blue", main = "Truncation_error by trapezoidal rule")
```

```{r}
# Truncation error of Simpson's rule
N <- 1e4
a <- seq(1e2, 1e4, by = 45)
Truncation_error <- sapply(a, TE_simp, n = N) 
plot(a, Truncation_error, type = "l", col = "red", main = "Truncation_error by simpson's rule")
```
For both rules, the truncation error decrease  as a gets larger and larger and fluctuate around 0. 

### c)
```{r}
steps_trape <- function(a) {
    diff <- abs(trapezoidal(a, 102)-trapezoidal(a, 101))
    n0 = 103
    n = n0
    epsilonI <- 1e-4
    while(diff > epsilonI) {
        diff <- abs(trapezoidal(a, n)-trapezoidal(a, n-1))
        n <- n + 1
    }
    count <- n - n0 +1
    return(count)
}

steps_simp <- function(a) {
    diff <- abs(simpson(a, 102)-simpson(a, 101))
    n0 = 103
    n = n0
    epsilonI <- 1e-4
    while(diff > epsilonI) {
        diff <- abs(simpson(a, n)-simpson(a, n-1))
        n <- n + 1
    }
    count <- n - n0 +1
    return(count)
}
```

```{r}
a <- seq(50, 1000, by=10)
steps_by_tr <- sapply(a, steps_trape)
steps_by_si <- sapply(a, steps_simp)
plot(a, steps_by_si, type = "l", col = "blue",  ylab = "steps")
lines(a, steps_by_tr, col = "red")
legend("topright", c("trapezoidal", "simpson"),
       col=c("red", "blue"))
```
When a increases, steps processes by simpson's rule also increase but can still converge when steps are large enough. While steps by trapezoidal rule will converge to a number around 80 in this case.

### d)
For another integrand:
```{r}
fx2 <- function(x) {
    f <- 1 + exp(-x)*sin(8*x^(2/3))
}
```

```{r}
trapezoidal2 <- function(n) {
    a1 <- 0
    a2 <- 2
    x <- seq(a1, a2, length.out = n)
    h <- x[2] - x[1]
    wn <- rep(h, n)
    wn[1] <- h/2
    wn[n] <- h/2
    fn <- sapply(x, fx2)
    ans <- sum(fn*wn)
    return(ans)
} 

simpson2 <- function(n) {
    a1 <- 0
    a2 <- 2
    x <- seq(a1, a2, length.out = n)
    h <- x[2] - x[1]
    wn <- rep(c(0, 4*h/3), floor(n/2)) + rep(c(2*h/3, 0), floor(n/2))
    wn[1] <- h/3
    wn[n] <- h/3
    fn <- sapply(x, fx2)
    ans <- sum(fn*wn)
    return(ans)
} 

epsilon <- 1
N <- 11
I1 <- trapezoidal2(10)
while (epsilon > 1e-4) {
    I2 <- trapezoidal2(N)
    epsilon <- abs(I1 - I2)
    I1 <- I2
    N <- N + 1
}
print(I1)

epsilon <- 1
N <- 11
I1 <- simpson2(10)
while (epsilon > 1e-4) {
    I2 <- simpson2(N)
    epsilon <- abs(I1 - I2)
    I1 <- I2
    N <- N + 1
}
print(I1)
```
The integral by trapezoidal rule converges to 2.0126 while the integral by simpson's rule converges to 2.0162. These two integral are very close.

## Question 3
```{r}
S0 <- 1
TT <- 5
t <- 0
tau <- TT-t
r <- 0
V0 <- 0.1
theta <- 0.1
sigma <- 0.2
rho <- -0.3
lambda <- 0
q <- 0 # Divident yield

a <- 1000
N <- 1001 # must be an integer

# The Simpson’s rule
simpsonReal <- function(n, a, i, K, kappa) {
    x <- seq(1/a, a, length.out = n)
    h <- x[2] - x[1]
    wn <- rep(c(2*h/3, 4*h/3), floor(n/2))
    wn[1] <- h/3
    wn[n] <- h/3
    fn <- sapply(x, RE, i = i, K = K, kappa = kappa)
    ans <- sum(fn*wn)
    return(ans)
}

ub <- function(i, kappa) {
    if (i == 1) {
        u <- 0.5
        b <- kappa + lambda - rho*sigma
    }
    else if (i == 2) {
        u <- -0.5
        b <- kappa + lambda
    }
    ub <- c(u, b)
    return(ub)
}

C <- function(x, i, kappa) {
    u <- ub(i, kappa)[1]
    b <- ub(i, kappa)[2]
    d <- sqrt((rho*sigma*x*1i - b)^2 - sigma^2*(2*u*x*1i - x^2))
    g <- (b - rho*sigma*x*1i + d)/(b - rho*sigma*x*1i - d)
    cc <- (r - q)*x*tau*1i + 
        kappa*theta/sigma^2*((b - rho*sigma*x*1i + d)*tau - 2*log((1 - g*exp(d*tau))/(1 - g)))
    return(cc)
}

D <- function(x, i, kappa) {
    u <- ub(i, kappa)[1]
    b <- ub(i, kappa)[2]
    d <- sqrt((rho*sigma*x*1i - b)^2 - sigma^2*(2*u*x*1i - x^2))
    g <- (b - rho*sigma*x*1i + d)/(b - rho*sigma*x*1i - d)
    dd <- (b - rho*sigma*x*1i + d)/sigma^2*(1 - exp(d*tau))/(1 - g*exp(d*tau))
    return(dd)
}

phi <- function(x, i, kappa) {
    phi <- exp(C(x, i, kappa) + D(x, i, kappa)*V0 + 1i*x*log(S0))
    return(phi)
}

RE <- function(x, i, K, kappa) {
    temp <- exp(-1i*x*log(K))*phi(x, i, kappa)/(1i*x)
    RE <- Re(temp)
    RE[is.nan(RE)] = 0
    return(RE)
}

Calloption <- function(N, a, K, kappa) {
    P1 <- 1/2 + 1/pi*simpsonReal(N, a, 1, K, kappa)
    P2 <- 1/2 + 1/pi*simpsonReal(N, a, 2, K, kappa)
    C <- S0*P1 - K*exp(-(r-q)*tau)*P2
    return(C)
}

K <- c(0.5, 0.75, 1, 1.25, 1.5)
kappa <- c(1, 2, 4)

temp <- as.data.frame(matrix(rep(0, 15), nrow = 5))
comparetable <- data.frame(temp)
rownames(comparetable) <- as.character(K)

for (i in 1:5) {
    for (j in 1:3) {
        k <- K[i]
        kap <- kappa[j]
        Call <- Calloption(N, a, k, kap)
        comparetable[i, j] <- Call
    }
}
Real_value <- c(0.543017, 0.385109, 0.273303, 0.195434, 0.14121)
comparetable <- data.frame(comparetable, Real_value)
colnames(comparetable)[1:3] <- c("kappa=1","kappa=2","kappa=4")

```

The result is showed below:

```{r results = 'asis'}
knitr::kable(comparetable, caption = "Table")
```

