---
title: "621_Homework 2"
output:
  pdf_document:
    toc: true
    toc_depth: 3
author: "Liting Hu"
---
```{r global_options, include=FALSE}
knitr::opts_chunk$set(warning=FALSE, message=FALSE)
```

```{r message = FALSE, echo = FALSE}
library(quantmod)
library(lubridate)
library(ggplot2)
library(cowplot)
```

## Question 1
### a)
Construct function to calculate option values using an additive binomial tree:
```{r}
Option_BT <- function(isCall=T, isAmerican=F, S0=100, K=100, 
                       Tm=1, sigma=0.2, r=0.0075, N=500) {
    # The additive binomial tree
    dt <- Tm/N
    nu <- r - 0.5*sigma^2
    dx <- sqrt(sigma^2*dt + (nu*dt)^2)
    p <- 0.5 + 0.5*(nu*dt/dx)
    
    disc <- exp(-r*dt)
    St <- rep(0, N + 1)
    C <- rep(0, N + 1)
    P <- rep(0, N + 1)
    St[1] <- S0*exp(-N*dx)
    for (i in 1:N) {St[i+1] <- St[i]*exp(2*dx)}
    C <- pmax(C, St - K)
    P <- pmax(P, K - St)
    for (i in N:1) {
        for (j in 1:i) {
            C[j] = disc*(p*C[j + 1] + (1 - p)*C[j])
            P[j] = disc*(p*P[j + 1] + (1 - p)*P[j])
            St[j] <- St[j]/exp(-dx)
            if (isAmerican==T) {
                C[j] = max(C[j], St[j] - K)
                P[j] = max(P[j], K - St[j])}
        }
    }
    if (isCall) ans <- C[1]
    else ans <- P[1]
    return(ans)
}

```

### b)
This is BSM formula:
```{r}
Option_BSM <- function(isCall = T, S0=100, K=100, Tm=1, sigma=0.2, r=0.0075, div=0) {
    d1 <- (log(S0/K) + (r - div + sigma^2/2)*Tm)/sigma/sqrt(Tm)
    d2 <- d1 - sigma*sqrt(Tm)
    if (isCall) {p <- S0*exp(-div*Tm)*pnorm(d1) - K*exp(-r*Tm)*pnorm(d2)}
    else {p <- K*exp(-r*Tm)*pnorm(-d2) - S0*exp(-div*Tm)*pnorm(-d1)}
    return(p)
}
```

Then download options data and stock data from yahoo finance:
```{r}
setwd("/Users/apple/Desktop/621/HW2")
readfile <- try(read.csv("Calls.csv"), TRUE)
if (inherits(readfile, "try-error")) {
    # Get option chains from Yahoo finance
    AAPL.OPTS <- getOptionChain("AAPL", NULL)
    C1 <- AAPL.OPTS$Apr.21.2017$calls
    P1 <- AAPL.OPTS$Apr.21.2017$puts
    C1$Ave.Price <- (C1$Bid + C1$Ask)/2
    P1$Ave.Price <- (P1$Bid + P1$Ask)/2
    C1 <- C1[, c(1, 4, 5, 8)]
    P1 <- P1[, c(1, 4, 5, 8)]
    C2 <- AAPL.OPTS$May.19.2017$calls
    P2 <- AAPL.OPTS$May.19.2017$puts
    C2$Ave.Price <- (C2$Bid + C2$Ask)/2
    P2$Ave.Price <- (P2$Bid + P2$Ask)/2
    C2 <- C2[, c(1, 4, 5, 8)]
    P2 <- P2[, c(1, 4, 5, 8)]
    C3 <- AAPL.OPTS$Jun.16.2017$calls
    P3 <- AAPL.OPTS$Jun.16.2017$puts
    C3$Ave.Price <- (C3$Bid + C3$Ask)/2
    P3$Ave.Price <- (P3$Bid + P3$Ask)/2
    C3 <- C3[, c(1, 4, 5, 8)]
    P3 <- P3[, c(1, 4, 5, 8)]
    
    temp <- merge(C1, C2, by = "Strike")
    calls <- merge(temp, C3, by = "Strike")
    temp <- merge(P1, P2, by = "Strike")
    puts <- merge(temp, P3, by = "Strike")
    coln <- c("Strike", "Bid1", "Ask1", "Apr.21.2017", "Bid2", "Ask2", 
              "May.19.2017", "Bid3", "Ask3", "Jun.16.2017")
    colnames(calls) <- coln
    colnames(puts) <- coln
    calls <- calls[3:22, ]
    puts <- puts[-19, ]
    print(calls)
    print(puts)
    write.csv(calls, file = "Calls.csv", row.names = F)
    write.csv(puts, file = "Puts.csv", row.names = F)
} else {
    calls <- read.csv("Calls.csv")
    puts <- read.csv("Puts.csv")
}

todaystock <- getQuote("AAPL") 
# S_0 <- todaystock[, 2] # The value of underlying
# When the option data are downloaded, the stock price is 139.24.
S_0 <- 139.55
```

As what we do in HW1, calculate the implied vols for all strikes and all maturities. I will skip this part but just show the result.
```{r echo=FALSE}
tau <- c(35/252, 55/252, 73/252) # time to maturity
r <- 0.0075 # the current short-term interest rate
fsigma <- function(isCall = T, sigma, K_i, maturity_i) {
    # Epsilon. To calculate implied vol
    S0 <- S_0
    if (isCall) {prices <- calls}
    else {prices <- puts}
    p <- prices[K_i, 3*maturity_i + 1]
    K <- calls[K_i, 1]
    price.by.bs <- Option_BSM(isCall, S0, K, tau[maturity_i], sigma, r)
    ans <- price.by.bs - p
    return(ans)
}

IV.Calls <- calls[, c(1, 4, 7, 10)]
IV.PUTS <- puts[, c(1, 4, 7, 10)]
for(i in 1:20) {
    for(j in 1:3) {
        # use bisection method to calculate implied vols
        a <- 0.001
        b <- 100
        epsilon <- abs(a - b)
        while(epsilon > 1e-4) {
            mid <- (a + b)/2
            if (fsigma(T, a, i, j)*fsigma(T, mid, i, j) < 0 ) b <- mid
            else a <- mid
            epsilon <- abs(a - b)
        }
        IV.Calls[i, j+1] <- a
        a <- 0.001
        b <- 100
        epsilon <- abs(a - b)
        while(epsilon > 1e-4) {
            mid <- (a + b)/2
            if (fsigma(F, a, i, j)*fsigma(F, mid, i, j) < 0 ) b <- mid
            else a <- mid
            epsilon <- abs(a - b)
        }
        IV.PUTS[i, j+1] <- a
    }
}
IV.Calls[,2:4][IV.Calls[,2:4]>99]=NaN
IV.PUTS[,2:4][IV.PUTS[,2:4]>99]=NaN # Implied vols calculated as in HW1
```

Implied vols are showed below:

```{r results = 'asis'}
knitr::kable(IV.Calls, caption = "Implied vols for Call Option")
```

```{r results = 'asis'}
knitr::kable(IV.PUTS, caption = "Implied vols for Put Option")
```
P.S. Sorry about losing put options data. When I noticed that I could just donwload them beyond trading time.


Using these vols, calculate option prices by Binomial Tree and BSM:
```{r}
Calls.Prices <- merge(IV.Calls, IV.Calls, by = "Strike") # Define data frames for options prices
Puts.Prices <- merge(IV.PUTS, IV.PUTS, by = "Strike")
coln2 <- c("Strike", "Apr.21.2017.BT", "Apr.21.2017.BSM", "May.19.2017.BT", 
           "May.19.2017.BSM", "Jun.16.2017.BT", "Jun.16.2017.BSM") 
colnames(Calls.Prices) <- coln2
colnames(Puts.Prices) <- coln2
coln3 <- c("Strike", "Apr.21.2017.A", "May.19.2017.A", "Jun.16.2017.A") 
Calls.A <- IV.Calls # for American Options
Puts.A <- IV.PUTS
colnames(Calls.A) <- coln3
colnames(Puts.A) <- coln3

Strike.C <- calls[, 1]
Strike.P <- puts[, 1]

for(i in 1:20) {
    for(j in 1:3) {
        # Calculate option prices
        sigma.C <- IV.Calls[i, j+1]
        sigma.P <- IV.PUTS[i, j+1]
        Calls.Prices[i, 2*j] <- Option_BT(T, F, S_0, Strike.C[i], tau[j], sigma.C, r, 200)
        Calls.Prices[i, 2*j+1] <- Option_BSM(T, S_0, Strike.C[i], tau[j], sigma.C, r)
        Puts.Prices[i, 2*j] <- Option_BT(F, F, S_0, Strike.C[i], tau[j], sigma.P, r, 200)
        Puts.Prices[i, 2*j+1] <- Option_BSM(F, S_0, Strike.C[i], tau[j], sigma.P, r)
        Calls.A[i, j+1] <- Option_BT(T, T, S_0, Strike.C[i], tau[j], sigma.C, r, 200)
        Puts.A[i, j+1] <- Option_BT(F, T, S_0, Strike.C[i], tau[j], sigma.P, r, 200)
    }
}
```
Option prices calculated as below:
```{r results = 'asis'}
knitr::kable(Calls.Prices, caption = "Call option prices")
```

```{r results = 'asis'}
knitr::kable(Puts.Prices, caption = "Put option prices")
```

### c)
```{r echo=F}
mergedf.c <- merge(calls, Calls.Prices, by = "Strike")
mergedf.c <- merge(mergedf.c, Calls.A, by = "Strike")
mergedf.p <- merge(puts, Puts.Prices, by = "Strike")
mergedf.p <- merge(mergedf.p, Puts.A, by = "Strike")

C.A <- ggplot(data = mergedf.c, aes(x = Strike)) + 
    geom_line(aes(y=Apr.21.2017.A, colour = "red")) +
    geom_line(aes(y=May.19.2017.A, colour = "green")) + 
    geom_line(aes(y=Jun.16.2017.A, colour = "blue")) +
    geom_line(aes(y=Bid1), colour = "red", linetype=2) +
    geom_line(aes(y=Ask1), colour = "red", linetype=3) +
    geom_line(aes(y=Bid2), colour = "green", linetype=2) +
    geom_line(aes(y=Ask2), colour = "green", linetype=3) +
    geom_line(aes(y=Bid3), colour = "blue", linetype=2) +
    geom_line(aes(y=Ask3), colour = "blue", linetype=3) +
    scale_colour_manual("Maturity", breaks = c("blue", "green", "red"), 
                        values = c("blue", "green", "red"), labels=c("Jun.16", "May.19", "Apr.21")) +
    ylab("Price") + 
    labs(title="American Calls by Binomial Tree")
C.BT <- ggplot(data = mergedf.c, aes(x = Strike)) + 
    geom_line(aes(y=Apr.21.2017.BT, colour = "red")) +
    geom_line(aes(y=May.19.2017.BT, colour = "green")) + 
    geom_line(aes(y=Jun.16.2017.BT, colour = "blue")) +
    geom_line(aes(y=Bid1), colour = "red", linetype=2) +
    geom_line(aes(y=Ask1), colour = "red", linetype=3) +
    geom_line(aes(y=Bid2), colour = "green", linetype=2) +
    geom_line(aes(y=Ask2), colour = "green", linetype=3) +
    geom_line(aes(y=Bid3), colour = "blue", linetype=2) +
    geom_line(aes(y=Ask3), colour = "blue", linetype=3) +
    scale_colour_manual("Maturity", breaks = c("blue", "green", "red"), 
                        values = c("blue", "green", "red"), labels=c("Jun.16", "May.19", "Apr.21")) +
    ylab("Price") + 
    labs(title="European Calls by Binomial Tree")
C.BSM <- ggplot(data = mergedf.c, aes(x = Strike)) + 
    geom_line(aes(y=Apr.21.2017.BSM, colour = "red")) +
    geom_line(aes(y=May.19.2017.BSM, colour = "green")) + 
    geom_line(aes(y=Jun.16.2017.BSM, colour = "blue")) +
    geom_line(aes(y=Bid1), colour = "red", linetype=2) +
    geom_line(aes(y=Ask1), colour = "red", linetype=3) +
    geom_line(aes(y=Bid2), colour = "green", linetype=2) +
    geom_line(aes(y=Ask2), colour = "green", linetype=3) +
    geom_line(aes(y=Bid3), colour = "blue", linetype=2) +
    geom_line(aes(y=Ask3), colour = "blue", linetype=3) +
    scale_colour_manual("Maturity", breaks = c("blue", "green", "red"), 
                        values = c("blue", "green", "red"), labels=c("Jun.16", "May.19", "Apr.21")) +
    ylab("Price") + 
    labs(title="European Calls by BSM")

P.A <- ggplot(data = mergedf.p, aes(x = Strike)) + 
    geom_line(aes(y=Apr.21.2017.A, colour = "red")) +
    geom_line(aes(y=May.19.2017.A, colour = "green")) + 
    geom_line(aes(y=Jun.16.2017.A, colour = "blue")) +
    geom_line(aes(y=Bid1), colour = "red", linetype=2) +
    geom_line(aes(y=Ask1), colour = "red", linetype=3) +
    geom_line(aes(y=Bid2), colour = "green", linetype=2) +
    geom_line(aes(y=Ask2), colour = "green", linetype=3) +
    geom_line(aes(y=Bid3), colour = "blue", linetype=2) +
    geom_line(aes(y=Ask3), colour = "blue", linetype=3) +
    scale_colour_manual("Maturity", breaks = c("blue", "green", "red"), 
                        values = c("blue", "green", "red"), labels=c("Jun.16", "May.19", "Apr.21")) +
    ylab("Price") + 
    labs(title="American Puts by Binomial Tree")
P.BT <- ggplot(data = mergedf.p, aes(x = Strike)) + 
    geom_line(aes(y=Apr.21.2017.BT, colour = "red")) +
    geom_line(aes(y=May.19.2017.BT, colour = "green")) + 
    geom_line(aes(y=Jun.16.2017.BT, colour = "blue")) +
    geom_line(aes(y=Bid1), colour = "red", linetype=2) +
    geom_line(aes(y=Ask1), colour = "red", linetype=3) +
    geom_line(aes(y=Bid2), colour = "green", linetype=2) +
    geom_line(aes(y=Ask2), colour = "green", linetype=3) +
    geom_line(aes(y=Bid3), colour = "blue", linetype=2) +
    geom_line(aes(y=Ask3), colour = "blue", linetype=3) +
    scale_colour_manual("Maturity", breaks = c("blue", "green", "red"), 
                        values = c("blue", "green", "red"), labels=c("Jun.16", "May.19", "Apr.21")) +
    ylab("Price") + 
    labs(title="European Puts by Binomial Tree")
P.BSM <- ggplot(data = mergedf.p, aes(x = Strike)) + 
    geom_line(aes(y=Apr.21.2017.BSM, colour = "red")) +
    geom_line(aes(y=May.19.2017.BSM, colour = "green")) + 
    geom_line(aes(y=Jun.16.2017.BSM, colour = "blue")) +
    geom_line(aes(y=Bid1), colour = "red", linetype=2) +
    geom_line(aes(y=Ask1), colour = "red", linetype=3) +
    geom_line(aes(y=Bid2), colour = "green", linetype=2) +
    geom_line(aes(y=Ask2), colour = "green", linetype=3) +
    geom_line(aes(y=Bid3), colour = "blue", linetype=2) +
    geom_line(aes(y=Ask3), colour = "blue", linetype=3) +
    scale_colour_manual("Maturity", breaks = c("blue", "green", "red"), 
                        values = c("blue", "green", "red"), labels=c("Jun.16", "May.19", "Apr.21")) +
    ylab("Price") + 
    labs(title="European Puts by BSM")

plot_grid(C.A, C.BT, C.BSM, ncol = 1, nrow = 3)
```

\pagebreak
```{r echo=FALSE}
plot_grid(P.A, P.BT, P.BSM, ncol = 1, nrow = 3)
```

From charts and plots, results calculated by binomial tree and BSM model are similar, even with Amrican ones.


### d)
```{r}
Error.by.N <- function(N) {
    # error for European Put
    epsilon <- abs(Option_BSM(isCall = F) - Option_BT(isCall = F, N = N))
    return(epsilon)
} 

N <- c(10, 20, 30, 40, 50, 100, 150, 200, 250, 300, 350, 400)
Error <- sapply(N, Error.by.N)
plot(N, Error, type = "l")
```

From this plot we can see that as N goes up, the error is decreasing and approaching a very small number (near zero).

## Question 2
### a)

Implement a trinomial tree to price European, American Call and Put options:
```{r tt}
Option_TT <- function(isCall=T, isAmerican=F, S0=100, K=100, 
                      Tm=1, sigma=0.25, r=0.06, div=0.03, N=200) {
    # The Trinomial Tree
    dt <- Tm/N
    nu <- r - div - 0.5*sigma^2
    dx <- sigma*sqrt(3*dt)
    edx <- exp(dx)
    
    pu = 0.5*((sigma^2*dt + nu^2*dt^2)/dx^2 + nu*dt/dx)
    pm = 1.0 - (sigma^2*dt + nu^2*dt^2)/dx^2 
    pd = 0.5*((sigma^2*dt + nu^2*dt^2)/dx^2 - nu*dt/dx)
    
    disc <- exp(-r*dt) # discount factor
    St <- rep(0, 2*N + 1)
    Ct <- rep(0, 2*N + 1)
    Pt <- rep(0, 2*N + 1)
    St[1] <- S0*exp(-N*dx)
    for (i in 1:(2*N)) {St[i+1] <- St[i]*edx}
    Ct <- pmax(Ct, St - K)
    Pt <- pmax(Pt, K - St)
    
    temp <- matrix(rep(0, (N+1)*(2*N+1)), ncol = (N + 1))
    C <- temp
    P <- temp
    C[, N+1] <- Ct
    P[, N+1] <- Pt
    
    for (i in N:1) {
        for (j in (N - i + 2):(N + i)) {
            C[j, i] = disc*(pu*C[j+1, i+1] + pm*C[j, i+1] + pd*C[j-1, i+1])
            P[j, i] = disc*(pu*P[j+1, i+1] + pm*P[j, i+1] + pd*P[j-1, i+1])
            St[j] <- St[j]/exp(-dx)
            if (isAmerican==T) {
                C[j, i] = max(C[j, i], St[j] - K)
                P[j, i] = max(P[j, i], K - St[j])}
        }
    }
    if (isCall) {ans <- C[N + 1, 1]
    } else {ans <- P[N + 1, 1]}
    return(ans)
}
```


### b)
```{r tt2}
# European & American Calls
Call <- c(Option_TT(), Option_BSM(sigma = 0.25, r = 0.06, div = 0.03),
          Option_TT(isAmerican = T))

# European & American Calls Puts
Put <- c(Option_TT(isCall = F), Option_BSM(isCall = F, sigma = 0.25, r = 0.06, div = 0.03),
         Option_TT(isCall = F, isAmerican = T))
df <- data.frame(Call, Put)
rownames(df) <- c('Trinomial Tree', "Black-Scholes", "Trinomial Tree American")
```

```{r results = 'asis'}
knitr::kable(df, caption = "Pricing by trinomial tree")
```
Using given parameters, the results for European Call and Put are very close between trinomial tree method and BSM method. But when calculating American call option by trinomial tree, it doesn't converge

```{r}
Error.by.N <- function(N) {
    # error for European Put
    # for trinomial tree
    epsilon <- abs(Option_BSM(isCall = F, sigma = 0.25, r=0.06, div = 0.03) -
                       Option_TT(isCall = F, N = N))
    return(epsilon)
} 

N <- c(10, 20, 30, 40, 50, 100, 150, 200, 250, 300, 350, 400)
Error <- sapply(N, Error.by.N)
plot(N, Error, type = "l", col = "red")
```

Similar to binomial tree senario, from this plot we can see that as N goes up, the error is decreasing and approaching zero.

## Question 3
### a)
```{r}
Option_UPandOUT <- function(S0=10, K=10, Tm=0.3, sigma=0.2, r=0.01, div=0, H=11, N=500) {
    # European Up-and Out
    # Binomial tree
    dt <- Tm/N
    nu <- r - div - 0.5*sigma^2
    dx <- sqrt(sigma^2*dt + (nu*dt)^2)
    p <- 0.5 + 0.5*(nu*dt/dx)
    
    disc <- exp(-r*dt)
    St <- rep(0, N + 1)
    C <- rep(0, N + 1)
    St[1] <- S0*exp(-N*dx)
    for (i in 1:N) {St[i+1] <- St[i]*exp(2*dx)}
    C <- pmax(C, St - K)
    C[C >= H - K] = 0
    
    for (i in N:1) {
        for (j in 1:i) {
            St[j] <- St[j]/exp(-dx)
            if (St[j] >= H) {C[j] = 0
            } else {C[j] = disc*(p*C[j + 1] + (1 - p)*C[j])}
        }
    }
    ans <- C[1]
    return(ans)
}
Option_UPandOUT(N=10000)
```
The price of the European Up-and-Out call option is 0.0549897.

### b)
```{r}
Option_UPandOUT_BS <- function(S0=10, K=10, Tm=0.3, sigma=0.2, r=0.01, div=0, H=11) {
    # European Up-and-Out
    # Explicit formulas
    nu <- r - div - 0.5*sigma^2
    d_sh <- (log(S0/H) + nu*Tm)/sigma/sqrt(Tm)
    d_hs <- (log(H/S0) + nu*Tm)/sigma/sqrt(Tm)
    C_bs <- Option_BSM(T, S0, K, Tm, sigma, r, div)
    C_bs.h <- Option_BSM(T, S0, H, Tm, sigma, r, div)
    C_bs.hs <- Option_BSM(T, H^2/S0, K, Tm, sigma, r, div)
    C_bs.h.hs <- Option_BSM(T, H^2/S0, H, Tm, sigma, r, div)
    ans <- C_bs - C_bs.h - (H - K)*exp(-r*Tm)*pnorm(d_sh) -
        (H/S0)^(2*nu/sigma^2)*(C_bs.hs - C_bs.h.hs - (H - K)*exp(-r*Tm)*pnorm(d_hs))
    return(ans)
}
Option_UPandOUT_BS()
```
The price of European Up-and-Out is 0.053.

### c)

To price an European Up-and-In call option by analytical solution:

```{r}
# The analytical solution
Option_UPandIn <- function(S0=10, K=10, Tm=0.3, sigma=0.2, r=0.01, div=0, H=11) {
    nu <- r - div - 0.5*sigma^2
    d_sh <- (log(S0/H) + nu*Tm)/sigma/sqrt(Tm)
    d_hs <- (log(H/S0) + nu*Tm)/sigma/sqrt(Tm)
    C_bs.h <- Option_BSM(T, S0, H, Tm, sigma, r)
    P_bs.hs <- Option_BSM(F, H^2/S0, K, Tm, sigma, r)
    P_bs.h.hs <- Option_BSM(F, H^2/S0, H, Tm, sigma, r)
    ans <- C_bs.h + (H - K)*exp(-r*Tm)*pnorm(d_sh) +
        (H/S0)^(2*nu/sigma^2)*(P_bs.hs - P_bs.h.hs + (H - K)*exp(-r*Tm)*pnorm(-d_hs))
    return(ans)
}
Option_UPandIn()
```
The price of European Up-and-In is 0.398 by the analytical solution.

By The In-Out parity:
```{r}
# The In-Out parity
# UI_BS + UO_BS = C_BS
Option_BSM(T, 10, 10, 0.3, 0.2, 0.01) - Option_UPandOUT_BS()
```
The price of European Up-and-In is 0.398 by the In-Out parity which is nearly equal to the former one.

### d)
```{r}
American_UpandIn <- function(S0=10, K=10, Tm=0.3, sigma=0.2, r=0.01, div=0, H=11) {
    nu <- r - div - 0.5*sigma^2
    d_bs <- (log(H^2/S0/min(H, K)) + nu*Tm)/sigma/sqrt(Tm)
    d_sh <- (log(S0/H) + nu*Tm)/sigma/sqrt(Tm)
    C_bs <- Option_BSM(T, S0, K, Tm, sigma, r)
    C_bs.h <- Option_BSM(T, S0, H, Tm, sigma, r)
    P_bs.h.hs <- Option_BSM(F, H^2/S0, min(H, K), Tm, sigma, r, div)
    ans <- (H/S0)^(2*nu/sigma^2)*(P_bs.h.hs - (min(H, K) - K)*exp(-r*Tm)*pnorm(-d_bs))
    if (K > H) {ans <- ans + C_bs - C_bs.h - (H - K)*exp(-r*Tm)*pnorm(d_sh)}
    return(ans)
}
American_UpandIn()
```

