---
title: "621_Homework 3"
output:
  pdf_document:
    toc: true
    toc_depth: 3
author: "Liting Hu"
---
```{r global_options, include=FALSE}
knitr::opts_chunk$set(warning=FALSE, message=FALSE)
```

```{r message = FALSE, echo = FALSE}
library(quantmod)
library(ggplot2)
library(reshape2)
```

## Question 1
### a)
Implement the Explicit Finite Difference method to price both European Call and Put options: 

```{r}
# European - Explicit Finite Difference method
Option_Ex <- function(isCall=T, S0=100, K=100, 
                      Tm=1, sigma=0.25, r=0.06, div=0.03, N=3, Nj=3) {
    # precompute constants
    dt <- Tm/N
    nu <- r - div - 0.5*sigma^2
    #dx <- 0.2
    dx <- sigma*sqrt(3*dt)
    edx <- exp(dx)
    
    pu = 0.5*dt*((sigma/dx)^2 + nu/dx)
    pm = 1.0 - dt*(sigma/dx)^2 - r*dt
    pd = 0.5*dt*((sigma/dx)^2 - nu/dx)
    
    # initialise asset prices at maturity
    St <- seq(1,2*Nj+1)
    St <- S0*edx^(St-1-Nj)
    
    # initialise option values at maturity
    temp <- matrix(0, ncol = (N + 1), nrow = (2*Nj + 1))
    if (isCall) {
        C <- temp
        C[, N+1] <- pmax(C[, N+1], St - K)
        
        # step back 
        for (i in N:1) {
            for(j in 2:(2*Nj)) {
                C[j, i] <- pu*C[j+1, i+1] + pm*C[j, i+1] + pd*C[j-1, i+1]
            }
            C[1, i] <- C[2, i]
            C[2*Nj+1, i] <- C[2*Nj, i] + (St[2*Nj+1] - St[2*Nj])
        }
        ans <- C[Nj+1, 1]
    }
    else {
        P <- temp
        P[, N+1] <- pmax(P[, N+1], K - St)
        
        # step back 
        for (i in N:1) {
            for(j in 2:(2*Nj)) {
                P[j, i] <- pu*P[j+1, i+1] + pm*P[j, i+1] + pd*P[j-1, i+1]
            }
            P[2*Nj+1, i] <- P[2*Nj, i]
            P[1, i] <- P[2, i] + (St[2] - St[1])
        }
        ans <- P[Nj+1, 1]
    }
    return(ans)
}
```

### b)
Implement the Implicit Finite Difference method to price European Call and Put options:

```{r}
# European - Implicit Finite Difference method
Option_Im <- function(isCall=T, S0=100, K=100, 
                      Tm=1, sigma=0.25, r=0.06, div=0.03, N=3, Nj=3) {
    # precompute constants
    dt <- Tm/N
    nu <- r - div - 0.5*sigma^2
    #dx <- 0.2
    dx <- sigma*sqrt(3*dt)
    edx <- exp(dx)
    
    pu = -0.5*dt*((sigma/dx)^2 + nu/dx)
    pm = 1.0 + dt*(sigma/dx)^2 + r*dt
    pd = -0.5*dt*((sigma/dx)^2 - nu/dx)
    
    # initialise asset prices at maturity
    St <- seq(1, 2*Nj+1)
    St <- S0*edx^(St - 1 - Nj)

    # initialise option values at maturity
    C <- matrix(0, ncol = 2, nrow = (2*Nj + 1))
    if (isCall) {
        C[, 1] <- pmax(C[, 1], St - K)
    }
    else {
        C[, 1] <- pmax(C[, 1], K - St)
    }
    
    #computer derivative boundary condition
    lambda_L <- C[2, 1] - C[1, 1]
    lambda_U <- C[2*Nj+1, 1] - C[2*Nj, 1]
    
    #step back through lattice
    for(i in (N-1) : 0){
        C <- solve_imp_tridiagonal(C, pu, pm, pd, lambda_L, lambda_U, Nj)
        C[, 1] <- C[, 2]
    }
    
    ans <- C[Nj+1, 1]
    return(ans)
}

# Solve implicit tridiagonal system
solve_imp_tridiagonal <- function(C,pu,pm,pd,lambda_L,lambda_U,Nj) {
    #substitute boundary condition at j=-Nj into j = -Nj+1
    vj <- Nj + 1
    pmp <- c()
    pp <- c()
    pmp[-Nj+vj+1] <- pm + pd
    pp[-Nj+vj+1] <- C[-Nj+1+vj, 1] + pd*lambda_L
    
    #eliminate upper diagonal
    for (j in (-Nj+2):(Nj-1)) {
        pmp[j+vj] <- pm - pu*pd/pmp[j-1+vj]
        pp[j+vj] <- C[j+vj, 1] - pp[j-1+vj]*pd/pmp[j-1+vj]
    }
    
    #use bounary condition at j = Nj and equation at j = Nj-1
    C[Nj+vj, 2] <- (pp[Nj-1+vj] + pmp[Nj-1+vj]*lambda_U)/(pu + pmp[Nj-1+vj])
    C[Nj-1+vj, 2] <- C[Nj+vj, 2] - lambda_U
    
    #back substitution
    for( j in (Nj-2):-Nj){
        C[j+vj, 2] = (pp[j+vj] - pu*C[j+1+vj, 2])/pmp[j+vj]
    }
    return(C)
}
```

### c)

Assume \[ \epsilon  \approx \Delta {x^2} + \Delta t\]

Since the convergence condition is \[\Delta x \ge \sigma \sqrt {3\Delta t}\]

so we have
\[\Delta t \le \frac{{\Delta {x^2}}}{{3{\sigma ^2}}}\]
\[\varepsilon  \le \left( {1 + \frac{1}{{3{\sigma ^2}}}} \right)\Delta {x^2} < 0.001\]
\[\Delta x < \frac{1}{{10}}\sqrt {\frac{{3{\sigma ^2}}}{{(3{\sigma ^2} + 1)10}}} \]
From Les Clewlow and Chris Strickland, a reasonable range of asset price values at the maturity date of the option is three standard deviations either side of the mean. (\[{n_{SD}} = 3 \times 2 = 6\])
\[\Delta x = \frac{{{n_{SD}}\sigma \sqrt T }}{{2{N_j} + 1}} < \frac{1}{{10}}\sqrt {\frac{{3{\sigma ^2}}}{{10(3{\sigma ^2} + 1)}}} \]
\[{N_j} > 5{n_{SD}}\sigma \sqrt {\frac{{10(3{\sigma ^2} + 1)T}}{{3{\sigma ^2}}}}  - \frac{1}{2}\]
Under these conditon,
\[\Delta t \le \frac{{\Delta {x^2}}}{{3{\sigma ^2}}} = \frac{{0.001}}{{\left( {3{\sigma ^2} + 1} \right)}}\]
\[N = 1/\Delta t \ge 1000\left( {3{\sigma ^2} + 1} \right)\]

With parameters given by (d):
\[\begin{array}{l}
\Delta x = {\rm{0.01256562}}\\
{N_j} = 60\\
\Delta t = {\rm{0.0008421053}}\\
N = 1221
\end{array}\]


### d)

```{r}
epsilon <- 0.001
n <- 6
sigma <- 0.25
Nj <- ceiling(((n*sigma)/sqrt(0.001/(1+1/3/sigma^2))-1)/2)
Nj
N <- ceiling(3*((2*Nj+1)/n)^2)
N

Ex.C <- Option_Ex(N = N, Nj = Nj)
Ex.P <- Option_Ex(isCall = F, N = N, Nj = Nj)
Im.C <- Option_Im(N = N, Nj = Nj)
Im.P <- Option_Im(isCall = F, N = N, Nj = Nj)

re <- data.frame(c(Ex.C, Im.C), c(Ex.P, Im.P))
colnames(re) <- c("Call", "Put")
rownames(re) <- c("Explicit", "Implicit")

```


Prices under these four schemes are showed below. \(N\) is 60 and \(N_j\) is 1221, as calculated in part c.

```{r results = 'asis'}
knitr::kable(re, caption = "Finite Difference methods")
```


### e)
```{r}
# BSM model
Option_BSM <- function(isCall = T, S0=100, K=100, Tm=1, sigma=.25, r=0.06, div=0.03) {
    d1 <- (log(S0/K) + (r - div + sigma^2/2)*Tm)/sigma/sqrt(Tm)
    d2 <- d1 - sigma*sqrt(Tm)
    if (isCall) {p <- S0*exp(-div*Tm)*pnorm(d1) - K*exp(-r*Tm)*pnorm(d2)}
    else {p <- K*exp(-r*Tm)*pnorm(-d2) - S0*exp(-div*Tm)*pnorm(-d1)}
    return(p)
}

```

Iterative procedure:

```{r}
# Explicit-Call ----
Nj.count <- 10
dif <- 1
bsp <- Option_BSM()
while (dif > epsilon) {
    Nj.count <- Nj.count + 1
    N <- ceiling(3*((2*Nj.count+1)/n)^2)
    op <- Option_Ex(N = N, Nj = Nj.count)
    dif <-  abs(op - bsp)
}
Ex.C <- op
Ex.C.Nj <- Nj.count
Ex.C.N <- N

```

```{r}
# Implicit-Call ----
Nj.count <- 10
dif <- 1
while (dif > epsilon) {
    Nj.count <- Nj.count + 10
    # When Nj is too large (150+), the implicit fucntion runs very slow. 
    # So at first let the step length be 10
    N <- ceiling(3*((2*Nj.count+1)/n)^2)
    op <- Option_Im(N = N, Nj = Nj.count)
    dif <-  abs(op - bsp)
}
dif <- 1
Nj.count <- Nj.count - 10
while (dif > epsilon) {
    Nj.count <- Nj.count + 1
    N <- ceiling(3*((2*Nj.count+1)/n)^2)
    op <- Option_Im(N = N, Nj = Nj.count)
    dif <-  abs(op - bsp)
}
Im.C <- op
Im.C.Nj <- Nj.count
Im.C.N <- N
```

```{r}
# Explicit-Put ----
Nj.count <- 10
dif <- 1
bsp <- Option_BSM(isCall = F)
while (dif > epsilon) {
    Nj.count <- Nj.count + 1
    N <- ceiling(3*((2*Nj.count+1)/n)^2)
    op <- Option_Ex(isCall = F, N = N, Nj = Nj.count)
    dif <-  abs(op - bsp)
}
Ex.P <- op
Ex.P.Nj <- Nj.count
Ex.P.N <- N
```

```{r}
# Implicit-Put ----
Nj.count <- 10
dif <- 1
while (dif > epsilon) {
    Nj.count <- Nj.count + 10
    # the step length is 10 as the same reason above
    N <- ceiling(3*((2*Nj.count+1)/n)^2)
    op <- Option_Im(isCall = F, N = N, Nj = Nj.count)
    dif <- abs(op - bsp)
}
dif <- 1
Nj.count <- Nj.count - 10
while (dif > epsilon) {
    Nj.count <- Nj.count + 1
    N <- ceiling(3*((2*Nj.count+1)/n)^2)
    op <- Option_Im(isCall = F, N = N, Nj = Nj.count)
    dif <-  abs(op - bsp)
}
Im.P <- op
Im.P.Nj <- Nj.count
Im.P.N <- N
```

```{r}
result2 <- data.frame(c(Ex.C, Im.C), c(Ex.P, Im.P))
colnames(result2) <- c("Call", "Put")
rownames(result2) <- c("Explicit", "Implicit")

Nj <- c(Ex.C.Nj, Im.C.Nj, Ex.P.Nj, Im.P.Nj)
N <- c(Ex.C.N, Im.C.N, Ex.P.N, Im.P.N)
dt <- 1/Nj
dx <- 0.25*sqrt(3*dt)
steps <- data.frame(Nj, N, dt, dx)
rownames(steps) <- c("Explicit.Call", "Implicit.Call", "Explicit.Put", "Implicit.Put")
```
Results are showed below, all of them are close to BSM model.
```{r results = 'asis'}
knitr::kable(result2, caption = "Finite Difference methods")
```
N, Nj, dt, dx are silghtly different from theoretical answers.
```{r results = 'asis'}
knitr::kable(steps, caption = "Finite Difference methods")
```

### f)
```{r}
sigmas <- seq(0.05, 0.6, by = 0.02)
Tm <- 1
div <- 0.03
r <- 0.06
pu <- c()
pm <- c()
pd <- c()

for (s in sigmas) {
    Nj <- ceiling(((n*s)/sqrt(0.001/(1+1/3/s^2))-1)/2)
    N = ceiling(3*((2*Nj+1)/n)^2)
    dt <- Tm/N
    #dx <- s*sqrt(3*dt)
    dx <- 0.02
    nu <- r - div - 0.5*s^2
    pu = c(pu, -0.5*dt*((s/dx)^2 + nu/dx))
    pm = c(pm, 1.0 + dt*(s/dx)^2 + r*dt)
    pd = c(pd, -0.5*dt*((s/dx)^2 - nu/dx))
}

ps <- data.frame("sigma" = sigmas, pu, pm, pd)
ps.long <- melt(ps, id="sigma")
ggplot(data = ps.long, aes(x = sigma, y=value, colour=variable)) + geom_line()
```
pu and pd have a similar path which has a tendency to decrease from zero. While pm is increasing as sigma grows larger.

### g)
Implement the Crank-Nicolson Finite Difference method and price both European Call and Put options:

```{r}
# European - The Crank-Nicolson Finite Difference method
Option_CN <- function(isCall=T, S0=100, K=100, 
                      Tm=1, sigma=0.25, r=0.06, div=0.03, N=3, Nj=3) {
    # precompute constants
    dt <- Tm/N
    nu <- r - div - 0.5*sigma^2
    #dx <- 0.2
    dx <- sigma*sqrt(3*dt)
    edx <- exp(dx)
    
    pu = -0.25*dt*((sigma/dx)^2 + nu/dx)
    pm = 1.0 + 0.5*dt*(sigma/dx)^2 + 0.5*r*dt
    pd = -0.25*dt*((sigma/dx)^2 - nu/dx)
    
    # initialise asset prices at maturity
    St <- seq(1, 2*Nj+1)
    St <- S0*edx^(St - 1 - Nj)
    
    # initialise option values at maturity
    C <- matrix(0, ncol = 2, nrow = (2*Nj + 1))
    if (isCall) {
        C[, 1] <- pmax(C[, 1], St - K)
    }
    else {
        C[, 1] <- pmax(C[, 1], K - St)
    }
    
    # computer derivative boundary condition
    lambda_L <- C[2, 1] - C[1, 1]
    lambda_U <- C[2*Nj+1, 1] - C[2*Nj, 1]
    
    # step back through lattice
    for(i in (N-1) : 0){
        C <- solve_CK_tridiagonal(C, pu, pm, pd, lambda_L, lambda_U, Nj)
        C[, 1] <- C[, 2]
    }
    return(C[Nj+1, 1])
}

# Solve CK tridiagonal system
solve_CK_tridiagonal <- function(C, pu, pm, pd, lambda_L, lambda_U, Nj){
    
    #substitute boundary condition at j=-Nj into j = -Nj+1
    vj <- Nj + 1
    pmp <- c(); pp <- c()
    pmp[-Nj+vj+1] <- pm + pd
    pp[-Nj+vj+1] <- (-pu*C[-Nj+2+vj, 1])-(pm-2)*C[-Nj+1+vj, 1] - pd*C[-Nj+vj, 1] + pd*lambda_L
    
    #eliminate upper diagonal
    for( j in(-Nj+2):(Nj-1)){
        pmp[j+vj] <- pm-pu*pd/pmp[j-1+vj]
        pp[j+vj] <- -pu*C[j+1+vj, 1] - (pm-2)*C[j+vj, 1]-pd*C[j-1+vj, 1]-pp[j-1+vj]*pd/pmp[j-1+vj]
    }
    
    #use bounary condition at j = Nj and equation at j = Nj-1
    C[Nj+vj, 2] <- (pp[Nj-1+vj] + pmp[Nj-1+vj]*lambda_U)/(pu + pmp[Nj-1+vj])
    C[Nj-1+vj, 2] <- C[Nj+vj, 2] - lambda_U
    
    #back substitution
    for( j in (Nj-2): (-Nj)){
        C[j+vj, 2] = ( pp[j+vj] - pu*C[j+1+vj, 2] )/pmp[j+vj]
    }
    C[1, 2] <- C[1, 1]-C[2, 1]+C[2, 2]
    return (C)
}

sigma <- 0.25
Nj <- ceiling(((n*sigma)/sqrt(0.001/(1+1/3/sigma^2))-1)/2)
N <- ceiling(3*((2*Nj+1)/n)^2)

CN.C <- Option_CN(N = N, Nj = Nj)
CN.P <- Option_CN(isCall = F, N = N, Nj = Nj)
result3 <- rbind(re, c(CN.C, CN.P))
rownames(result3) <- c("EFD", "IFD", "CNFD")
```

```{r results = 'asis'}
knitr::kable(result3, caption = "Finite Difference methods")
```

Prices by Crank-Nicolson are between what we calculate by explicit and implicit methods. All these three method have similar results.

### h)
```{r}
# calculate hedge sensitivities
Greeks <- function(isCall=T, S0=100, K=100, 
                   Tm=1, sigma=0.25, r=0.06, div=0.03, N=3, Nj=3) {
    # Output: Delta, Gamma, Theta, and Vega
    
    # precompute constants
    dt <- Tm/N
    nu <- r - div - 0.5*sigma^2
    #dx <- 0.2
    dx <- sigma*sqrt(3*dt)
    edx <- exp(dx)
    
    pu = 0.5*dt*((sigma/dx)^2 + nu/dx)
    pm = 1.0 - dt*(sigma/dx)^2 - r*dt
    pd = 0.5*dt*((sigma/dx)^2 - nu/dx)
    
    dsigma <- 0.0001*sigma # to compute vega, define delta sigma as a small fraction 
    
    # initialise asset prices at maturity
    St <- seq(1, 2*Nj+1)
    St <- S0*edx^(St-1-Nj)
    
    # initialise option values at maturity
    temp <- matrix(0, ncol = (N + 1), nrow = (2*Nj + 1))
    if (isCall) {
        C <- temp
        C[, N+1] <- pmax(C[, N+1], St - K)
        
        # step back 
        for (i in N:1) {
            for(j in 2:(2*Nj)) {
                C[j, i] <- pu*C[j+1, i+1] + pm*C[j, i+1] + pd*C[j-1, i+1]
            }
            C[1, i] <- C[2, i]
            C[2*Nj+1, i] <- C[2*Nj, i] + (St[2*Nj+1] - St[2*Nj])
        }
    }
    else {
        P <- temp
        P[, N+1] <- pmax(P[, N+1], K - St)
        
        # step back 
        for (i in N:1) {
            for(j in 2:(2*Nj)) {
                P[j, i] <- pu*P[j+1, i+1] + pm*P[j, i+1] + pd*P[j-1, i+1]
            }
            P[2*Nj+1, i] <- P[2*Nj, i]
            P[1, i] <- P[2, i] + (St[2] - St[1])
        }
        C <- P
    } 
    
    delta <- (C[Nj+2, 1] - C[Nj, 1])/(St[Nj+2] - St[Nj])
    gamma <- 2*((C[Nj+2, 1] - C[Nj+1, 1])/(St[Nj+2] - St[Nj+1]) - 
                    (C[Nj+1, 1] - C[Nj, 1])/(St[Nj+1] - St[Nj]))/(St[Nj+2] - St[Nj])
    theta <- (C[Nj+1, 2] - C[Nj+1, 1])/dt
    vega <- (Option_Ex(isCall, S0, K, Tm, sigma + dsigma, r, div, N, Nj) - 
        Option_Ex(isCall, S0, K, Tm, sigma - dsigma, r, div, N, Nj))/2/dsigma
    result <- c(delta, gamma, theta, vega)
    return(result)
}
```
Using parameters in (d), for a European option: Delta, Gamma, Theta, and Vega are 
```{r}
Greeks(N=N, Nj=Nj)
```

## Question 2
### a)
Doing same thing as in HW1:
```{r}
# Download option prices
setwd("/Users/apple/Desktop/621/HW3")
readfile <- try(read.csv("Calls.csv"), TRUE)
if (inherits(readfile, "try-error")) {
    # Get option chains from Yahoo finance
    AAPL.OPTS <- getOptionChain("AAPL", NULL)
    C1 <- AAPL.OPTS$Apr.21.2017$calls
    P1 <- AAPL.OPTS$Apr.21.2017$puts
    C1$Ave.Price <- (C1$Bid + C1$Ask)/2
    P1$Ave.Price <- (P1$Bid + P1$Ask)/2
    C1 <- C1[, c(1, 4, 5, 8)]
    P1 <- P1[, c(1, 4, 5, 8)]
    C2 <- AAPL.OPTS$May.19.2017$calls
    P2 <- AAPL.OPTS$May.19.2017$puts
    C2$Ave.Price <- (C2$Bid + C2$Ask)/2
    P2$Ave.Price <- (P2$Bid + P2$Ask)/2
    C2 <- C2[, c(1, 4, 5, 8)]
    P2 <- P2[, c(1, 4, 5, 8)]
    C3 <- AAPL.OPTS$Jun.16.2017$calls
    P3 <- AAPL.OPTS$Jun.16.2017$puts
    C3$Ave.Price <- (C3$Bid + C3$Ask)/2
    P3$Ave.Price <- (P3$Bid + P3$Ask)/2
    C3 <- C3[, c(1, 4, 5, 8)]
    P3 <- P3[, c(1, 4, 5, 8)]
    
    temp <- merge(C1, C2, by = "Strike")
    calls <- merge(temp, C3, by = "Strike")
    temp <- merge(P1, P2, by = "Strike")
    puts <- merge(temp, P3, by = "Strike")
    coln <- c("Strike", "Bid1", "Ask1", "Apr.21.2017", "Bid2", "Ask2", 
              "May.19.2017", "Bid3", "Ask3", "Jun.16.2017")
    colnames(calls) <- coln
    colnames(puts) <- coln
    calls <- calls[9:18, ]
    puts <- puts[10:19, ]
    print(calls)
    print(puts)
    write.csv(calls, file = "Calls.csv", row.names = F)
    write.csv(puts, file = "Puts.csv", row.names = F)
} else {
    calls <- read.csv("Calls.csv")
    puts <- read.csv("Puts.csv")
}

# todaystock <- getQuote("AAPL") 
# S_0 <- todaystock[, 2] # The value of underlying
# When the option data are downloaded, the stock price is 140.26
S_0 <- 140.26

tau <- c(24/252, 44/252, 62/252) # time to maturity
r <- 0.0075 # the current short-term interest rate
fsigma <- function(isCall = T, sigma, K_i, maturity_i) {
    # Epsilon. To calculate implied vol
    S0 <- S_0
    if (isCall) {prices <- calls}
    else {prices <- puts}
    p <- prices[K_i, 3*maturity_i + 1]
    K <- calls[K_i, 1]
    price.by.bs <- Option_BSM(isCall, S0, K, tau[maturity_i], sigma, r)
    ans <- price.by.bs - p
    return(ans)
}

IV.Calls <- calls[, c(1, 4, 7, 10)]
IV.PUTS <- puts[, c(1, 4, 7, 10)]
for(i in 1:10) {
    for(j in 1:3) {
        # use bisection method to calculate implied vols
        a <- 0.001
        b <- 100
        epsilon <- abs(a - b)
        while(epsilon > 1e-4) {
            mid <- (a + b)/2
            if (fsigma(T, a, i, j)*fsigma(T, mid, i, j) < 0 ) b <- mid
            else a <- mid
            epsilon <- abs(a - b)
        }
        IV.Calls[i, j+1] <- a
        a <- 0.001
        b <- 100
        epsilon <- abs(a - b)
        while(epsilon > 1e-4) {
            mid <- (a + b)/2
            if (fsigma(F, a, i, j)*fsigma(F, mid, i, j) < 0 ) b <- mid
            else a <- mid
            epsilon <- abs(a - b)
        }
        IV.PUTS[i, j+1] <- a
    }
}
IV.Calls[,2:4][IV.Calls[,2:4]>99]=NaN
IV.PUTS[,2:4][IV.PUTS[,2:4]>99]=NaN # Implied vols calculated as in HW1
```

Implied vols are showed below:

```{r results = 'asis'}
knitr::kable(IV.Calls, caption = "Implied vols for Call Option")
```

```{r results = 'asis'}
knitr::kable(IV.PUTS, caption = "Implied vols for Put Option")
```

### b)
```{r}
FD.calls <- calls
FD.puts <- puts
for (i in 1:3) {
    Strike.C <- IV.Calls[, 1]
    Strike.P <- IV.PUTS[, 1]
    for (j in 1:10) {
        sigma.C <- IV.Calls[j, i+1]
        sigma.P <- IV.PUTS[j, i+1]
        
        # Theoretical N and Nj to make error less than 0.001
        Nj.C <- ceiling(((n*sigma.C)/sqrt(0.001/(1+1/3/sigma.C^2))-1)/2)
        N.C <- ceiling(3*((2*Nj.C+1)/n)^2)
        Nj.P <- ceiling(((n*sigma.P)/sqrt(0.001/(1+1/3/sigma.P^2))-1)/2)
        N.P <- ceiling(3*((2*Nj.P+1)/n)^2)

        # compute all things we need    
        if (is.na(sigma.C)) {
            FD.calls[j, (3*i-1):(3*i+1)] <- c(NaN, NaN, NaN)
        } else {
            FD.calls[j, 3*i-1] <- Option_Ex(S0 = S_0, K = Strike.C[j], Tm = tau[i], 
                                            sigma = sigma.C, div = 0, N = N.C, Nj = Nj.C)
            FD.calls[j, 3*i] <- Option_Im(S0 = S_0, K = Strike.C[j], Tm = tau[i], 
                                          sigma = sigma.C, div = 0, N = N.C, Nj = Nj.C)
            FD.calls[j, 3*i+1] <- Option_CN(S0 = S_0, K = Strike.C[j], Tm = tau[i], 
                                            sigma = sigma.C, div = 0, N = N.C, Nj = Nj.C)
        }
        
        if (is.na(sigma.P)) {
            FD.puts[j, (3*i-1):(3*i+1)] <- c(NaN, NaN, NaN)
        } else {
            FD.puts[j, 3*i-1] <- Option_Ex(isCall = F, S0 = S_0, K = Strike.P[j], Tm = tau[i], 
                                           sigma = sigma.P, div = 0, N = N.P, Nj = Nj.P)
            FD.puts[j, 3*i] <- Option_Im(isCall = F, S0 = S_0, K = Strike.P[j], Tm = tau[i], 
                                         sigma = sigma.P, div = 0, N = N.P, Nj = Nj.P)
            FD.puts[j, 3*i+1] <- Option_CN(isCall = F, S0 = S_0, K = Strike.P[j], Tm = tau[i], 
                                           sigma = sigma.P, div = 0, N = N.P, Nj = Nj.P)
        }
    }
}

coln11 <- c("Strikes", "EFD1", "IFD1", "CNFD1", 
            "EFD2", "IFD2", "CNFD2", 
            "EFD3", "IFD3", "CNFD3")
# "1": maturity at Apr.21.2017; "2": maturity at May.19.2017; "3": maturity at Jun.16.2017.
colnames(FD.calls) <- coln11
colnames(FD.puts) <- coln11
```

All prices are showed below. For call and put options with same parameters, EFD < CNFD < IFD. 
```{r results = 'asis'}
knitr::kable(FD.calls, caption = "Prices of Call Option under Finite Difference methods")
```

```{r results = 'asis'}
knitr::kable(FD.puts, caption = "Prices of Put Option under Finite Difference methods")
```

### c)
```{r}
G.calls <- c()
G.puts <- c()
for (i in 1:10) {
    Strike.C <- IV.Calls[i, 1]
    Strike.P <- IV.PUTS[i, 1]
    G.c <- c()
    G.p <- c()
    for(j in 1:3) {
        sigma.C <- IV.Calls[i, j+1]
        sigma.P <- IV.PUTS[i, j+1]
        
        # Theoretical N and Nj to make error less than 0.001
        Nj.C <- ceiling(((n*sigma.C)/sqrt(0.001/(1+1/3/sigma.C^2))-1)/2)
        N.C <- ceiling(3*((2*Nj.C+1)/n)^2)
        Nj.P <- ceiling(((n*sigma.P)/sqrt(0.001/(1+1/3/sigma.P^2))-1)/2)
        N.P <- ceiling(3*((2*Nj.P+1)/n)^2)
        
        if (is.na(sigma.C)) {
            G.c <- c(G.c, NaN, NaN, NaN, NaN)
        } else {
            G.c <- c(G.c, Greeks(T, S_0, Strike.C, tau[j], sigma.C, r, 0, N.C, Nj.C))
        }
        
        if (is.na(sigma.P)) {
            G.p <- c(G.p, NaN, NaN, NaN, NaN)
        } else {
            G.p <- c(G.p, Greeks(F, S_0, Strike.P, tau[j], sigma.P, r, 0, N.P, Nj.P))
        }
    }
    G.calls <- rbind(G.calls, G.c)
    G.puts <- rbind(G.puts, G.p)
}
coln22 <- c("Delta1", "Gamma1", "Theta1", "Vega1", 
            "Delta2", "Gamma2", "Theta2", "Vega2", 
            "Delta3", "Gamma3", "Theta3", "Vega3")
# "1": maturity at Apr.21.2017; "2": maturity at May.19.2017; "3": maturity at Jun.16.2017.

colnames(G.calls) <- coln22
colnames(G.puts) <- coln22

G.calls <- cbind(calls[1], G.calls)
G.puts <- cbind(puts[1], G.puts)
```
Delta, Gamma, Theta, and Vega are showed below. 
P.S. "1": maturity at Apr.21.2017; "2": maturity at May.19.2017; "3": maturity at Jun.16.2017.
```{r results = 'asis'}
knitr::kable(G.calls, caption = "Call option greeks by Explicit Finite Difference methods")
```

```{r results = 'asis'}
knitr::kable(G.puts, caption = "Put option greeks by Explicit Finite Difference methods")
```

### d)
```{r}
a.big.df <- as.data.frame(matrix(rep(0, 10), nrow = 1))
coln33 <- c("T", "Strike", "Type", "Ask", "Bid", 
            "Market", "implied vol", "EFD", "IFD", "CNFD")
colnames(a.big.df) <- coln33
for (i in 1:3) {
    df1 <- cbind(tau[i], calls[, 1], "Call", calls[, (3*i-1):(3*i+1)], 
                 IV.Calls[, i+1], FD.calls[, (3*i-1):(3*i+1)])
    colnames(df1) <- coln33
    df2 <- cbind(tau[i], puts[, 1], "Put", puts[, (3*i-1):(3*i+1)], 
                 IV.PUTS[, i+1], FD.puts[, (3*i-1):(3*i+1)])
    colnames(df2) <- coln33
    a.big.df <- rbind(a.big.df, df1, df2)
}
a.big.df <- a.big.df[-1, ]
```

```{r results = 'asis'}
knitr::kable(a.big.df)
```

```{r}
Captions <- c("Call options mature at Apr.21.2017", "Put options mature at Apr.21.2017",
              "Call options mature at May.19.2017", "Put options mature at May.19.2017", 
              "Call options mature at Jun.16.2017", "Put options mature at Jun.16.2017")
for (i in 1:6) {
    df <- a.big.df[(10*i-9):(10*i), c(2, 4:6, 8:10)]
    df.long <- melt(df, id="Strike")
    print(ggplot(data=df.long, aes(x = Strike, y=value, colour=variable)) +
              geom_line() + ggtitle(Captions[i]))
    Sys.sleep(1)
}
```
All six kinds of prices coincide excepe CNFD. The reason might be that it needs more steps to convergence than explicit and implicit methods.

## Question 3
Define a function to calculate option price by heston model:
```{r}
Option_Heston <- function(S0=1, K=1, V0=0.1, theta=0.1, sigma=0.1, rou=-0.3, lambda=0, 
                          kappa=2, Tm=5, r=0, Nt=10000, Ns=40, Nj=40) {
    # Only for pricing European Put option in Heston model
    
    # Define maximum and minimun values for S, Nu, t
    Smin <- 0
    Smax <- 2*S0
    NUmin <- 0
    NUmax <- 0.5
    Tmin <- 0
    Tmax <- Tm
    
    # Building a uniform grid
    dS <- (Smax - Smin)/Ns
    St <- seq(Smin, Smax, by = dS)
    dnu <- (NUmax - NUmin)/Nj
    nu <- seq(NUmin, NUmax, by = dnu)
    dt <- (Tmax - Tmin)/Nt
    
    # initialise asset prices at maturity
    ST <- matrix(ncol = Nj + 1, nrow = Ns + 1)
    for (i in 1:(Ns+1)) {ST[i, ] <- St[i]}
    
    # initialise option values at maturity
    UT <- ST
    for (i in 1:(Ns+1)) {UT[i, ] <- pmax(K - UT[i, ], 0)}
    
    # At time t-dt
    UT_1 <- UT
    # step back
    for (k in 1:Nt) { 
        # by time
        for (i in 2:Ns) { 
            # by S
            for (j in 2:Nj) {
                # by nu
                
                # Based on the partial differential equation, calculate dU/dt
                # Then by dU/dt = ((U(t) - U(t-dt)))/dt, calcuate U(t-dt)
                dU.dS <- (UT[i+1, j] - UT[i-1, j])/2/dS
                dU.dnu <- (UT[i, j+1] - UT[i, j-1])/2/dnu
                dU.2dS <- (UT[i+1, j] - 2*UT[i, j] + UT[i-1, j])/dS^2
                dU.2dnu <- (UT[i, j+1] - 2*UT[i, j] + UT[i, j-1])/dnu^2
                dU.dS.dnu <- (UT[i+1, j+1] + UT[i-1, j-1] - UT[i+1, j-1] - UT[i-1, j+1])/4/dS/dnu
                
                dU.dt <- -(1/2*nu[j]*St[i]^2*dU.2dS + 1/2*sigma^2*nu[j]*dU.2dnu +
                               rou*sigma*nu[j]*St[i]*dU.dS.dnu + r*St[i]*dU.dS +
                               kappa*(theta - nu[j])*dU.dnu - r*UT[i, j])
                UT_1[i, j] <- UT[i, j] - dt*dU.dt
            }
        }
        
        # boundary conditon
        UT_1[1, ] <- UT_1[2, ] + dS # S=Smin
        UT_1[Ns+1, ] <- 0 # S=Smax
        
        # When nu = NUmax,  U(Si, NUmax, tn) = Si.
        # This column is identical in every step
        
        # When nu = NUmin
        for (l in 2:Ns) {
            dU.dS <- (UT[l+1, 1] - UT[l-1, 1])/2/dS
            dU.dnu <- (UT[l, 2] - UT[l, 1])/dnu
            dU.dt <- r*UT[l, 1] - r*St[l]*dU.dS - kappa*theta*dU.dnu
            UT_1[l, 1] <- UT[l, 1] - dt*dU.dt
        }
        
        UT <- UT_1
    }
    
    # Because the intervals in grid are small enough. 
    # we simply use the mean of surrounding grid points as our option price
    ii <- S0/dS + 1
    ii1 <- floor(ii)
    ii2 <- ii + 1
    jj <- V0/dnu + 1
    jj1 <- floor(jj)
    jj2 <- jj + 1
    ans <- (UT[ii1, jj1] + UT[ii1, jj2] + UT[ii2, jj1] + UT[ii2, jj2])/4
    return(ans)
}
```

```{r}
K <- c(0.5, 0.75, 1, 1.25, 1.5)
kappa <- c(1, 2, 4)

temp <- as.data.frame(matrix(ncol = 1, nrow = 5))
comparetable <- data.frame(temp)
rownames(comparetable) <- as.character(K)

for (i in 1:5) {
    kap = 2
    k <- K[i]
    PP <- Option_Heston(K = k, kappa = kap)
    comparetable[i, 1] <- PP
}
Real_value <- c(0.543017, 0.385109, 0.273303, 0.195434, 0.14121)
comparetable <- data.frame(comparetable, Real_value)
colnames(comparetable)[1] <- c("kappa=2") 
```

The result is showed below:

```{r results = 'asis'}
knitr::kable(comparetable, caption = "Comparison of solutions")
```